---
title: "COVID19"
author: "Bikram Maharjan"
date: "3/13/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

```{r, echo=FALSE}

library(tidyverse)
library(plotly)
library(reshape2)
library(data.table)


```



### Prepare the dataset

```{r}

# Read the daily originalCSV
daily <- read.csv('03-14-2020.csv', stringsAsFactors = FALSE)

# Read the time series data
covid_recovered <- read.csv('time_series_19-covid-Recovered.csv', stringsAsFactors = FALSE)
covid_confirmed <- read.csv('time_series_19-covid-Confirmed.csv', stringsAsFactors = FALSE)
covid_death <- read.csv('time_series_19-covid-Deaths.csv', stringsAsFactors = FALSE)

# The manually cleaned file [states_first_affected.csv]
states_confirmed_order <- read.csv('states_confirmed_order.csv')

# Sorted States starting with first known cases
states_confirmed_order_vec <- c(states_confirmed_order)

# Reorder the columns so that we can melt
daily <- daily[c(1,2,3,7,8,4,5,6)]

# Melt the reordered data
daily_long <- melt(daily, id=1:5)

# Prepare dataset for US only to get the first known case
usaDat_time_series <- covid_confirmed %>% filter(Country.Region == 'US') 

```


### Sort the US States from the first known confirmed cases
```{r}

# Initilize empty vector to store states from its first COVID-19 confirmed
states <- c()

# Melt the data so that we can count the first triggring 1 count instead of 0 count
usaDat_time_series_long <- melt(usaDat_time_series, id=1:4)

# US states / cities / county
length <- dim(usaDat_time_series_long)[1]

# Sort the U.S. State from its first appearances
for(i in 1:length){
  if(usaDat_time_series_long$value[i] > 0){
    positive_state <- usaDat_time_series_long$Province.State[i]
      if((positive_state %in% states)==FALSE){
      
        states[i] <- usaDat_time_series_long$Province.State[i]
  
      } 
  }
}

# Remove the NA and only get the subset
us_city_order <- subset(states, (!is.na(states)))

# Convert it into a DF
states_first_affected <- as.data.frame(us_city_order)

# Write to a file to do manual cleaning. 
# EG: removing city/county name and convert it into full State name
write.csv(states_first_affected, "states_first_affected.csv")

```

#### US City/State sorted from first known case
```{r}
# May have duplicate States but cleaning it manually
head(states_first_affected)
```


### U.S. only analysis
```{r}

# Prepare dataset for US only and plot using ggplotly
usaDat <- daily_long %>% filter(Country.Region == 'US') 

# Filter data with more than 0 cases
usaDat <- usaDat %>% filter(value > 0) 

# Order/Sort the DT using the first known State
usaDat_order <- setDT(usaDat)[order(match(Province.State, states_confirmed_order$states))]


# Prepare the plot
usaDat_plot <- ggplot(usaDat_order, aes(reorder(Province.State, -value), value, color = variable)) + 
                geom_point(stat = "identity") +
                ggtitle("COVID-19 cases in the US (As of March 14th 2020)") +
                xlab("U.S. States ") + 
                ylab("Total cases") +
                theme(axis.text.x = element_text(angle = 60, hjust = 1), 
                      plot.title = element_text(hjust = 0.5),
                      legend.title = element_text("Status of COVID-19"))

# plotly for US only states
ggplotly(usaDat_plot)

```


### Total Cases in the US
```{r}

# Prepare dataset for US only and plot using ggplotly
usaDat <- daily_long %>% filter(Country.Region == 'US') 

# Select the required vector
case_count <- usaDat[c(6,7)]

# Save the total case count group by cases [confirmed, recovered, death]
cases_count <- case_count %>% 
                type.convert(as.is = TRUE) %>% 
                group_by(variable) %>% 
                summarise_all(sum, na.rm =TRUE)

```

#### Total confirmed cases:
```{r}

confirmed_us <- cases_count %>% filter(variable == 'Confirmed')
print(confirmed_us$value)

```

#### Total recovered cases:
```{r}

recovered_us <- cases_count %>% filter(variable == 'Recovered')
print(recovered_us$value)

```

#### Total death cases:
```{r}

death_us <- cases_count %>% filter(variable == 'Deaths')
print(death_us$value)
```


### United States is the 9th most infected country as of March 14th, 2020 
#### not including China
```{r}

# List of distinct country of cases except `China`
except_china <- daily_long %>% filter(Country.Region != 'China') 

# Subset the required values
except_china_subset <- except_china[c(2,6,7)]

# Aggregrate the total distinct countries
except_chinaDat <- aggregate(value ~ except_china_subset$Country.Region + variable, except_china_subset, sum)

# Rename the columns with meaningful names for each
except_china_daily <- rename(except_chinaDat, country = c(1) , type_of_cases=c(2), count_total=c(3))

# show cases if it is more or equal to 100 confirmed cases
# More or equal to 5 deaths
# More or equal to 5 recovered
except_china_daily_100 <- except_china_daily %>% 
                            filter((type_of_cases == 'Confirmed' & count_total >= 100) |
                              (type_of_cases == 'Recovered' & count_total >= 5) | 
                              (type_of_cases == 'Deaths' & count_total >= 5) )


# Prepare the plot
plot <- ggplot(except_china_daily_100, aes(reorder(country, -count_total), count_total, color = type_of_cases)) + 
  geom_point(stat = "identity") +
  ggtitle("Countried with highest cases after China (As of March 14th 2020)")+
  xlab("World Country")+ 
  ylab("Total Cases")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1), 
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text("Status of COVID-19"))
  
```


#### Graph with more than 100 confirmed cases
#### Recovered and Deaths data are 5 or more cases
```{r}

# Graph with more than 100 confirmed cases - queal or more than 5 deaths and recovered
ggplotly(plot)

```



